---
title: "Assessing Normality Practice"
author: "Elise Haylett"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    code_folding: hide
    theme: yeti
    highlight: textmate
---

```{r globaloptions}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

```{r loadpackages}
library(tidyverse)
library(moments)
```

***
#### **Gas Mileage**

```{r}
mileage <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
```

```{r}
# Normality Assessment Here
mpg_xbar <- mean(mileage$mpg)
mpg_s <- sd(mileage$mpg)


# perfect normal distribution
ggplot(mileage, aes(x = mpg)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(mpg_xbar, mpg_s),   
                color = "purple") +
  scale_x_continuous(breaks = seq(mpg_xbar - (4 * mpg_s),  
                                  mpg_xbar + (4 * mpg_s), mpg_s))


# empirical cdf
ggplot(mileage, aes(x = mpg)) + 
  stat_ecdf() +
  stat_function(fun = pnorm, 
                args = list(mpg_xbar, mpg_s), 
                color = "purple")


# qqplot
ggplot(mileage, aes(sample = mpg)) + 
  stat_qq(alpha = .5) +
  stat_qq_line(color = "purple")


# skewness and kutosis
skewK_mpg <- tibble(statistic = c("Skewness", "Kurtosis", "Excess K"),
                theory = c(0, 3, 0),
                data   = round(c(skewness(mileage$mpg), 
                                 kurtosis(mileage$mpg), 
                                 kurtosis(mileage$mpg) - 3), 
                               2))
print.data.frame(skewK_mpg, row.names = FALSE)


```

* Looking at the normal curve --> Data is leptokurtic because the peak of the data density is taller than the peak of the theoretical normal curve (pointy)

* Stat_ecdf --> deviation from the normal S-curve * SHOW SOME SYSTEMATIC DEVIATION FROM NORMAL

* QQ plot --> does "tilde" thing in the middle * SHOW SOME SYSTEMATIC DEVIATION FROM NORMAL

* The skewness and kurtosis calculations confirm this. Skewenss is almost 0 (0.05) and it has some excess kurtosis (0/67)

***

#### **Old Faithful**

```{r}
geyser <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv")
```

```{r}
# Normality Assessment Here
ET_xbar <- mean(geyser$erupt_time)
ET_s <- sd(geyser$erupt_time)


# perfect normal distribution
ggplot(geyser, aes(x = erupt_time)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(ET_xbar, ET_s),   
                color = "orange") +
  scale_x_continuous(breaks = seq(ET_xbar - (4 * ET_s),  
                                  ET_xbar + (4 * ET_s), ET_s))


# empirical cdf
ggplot(geyser, aes(x = erupt_time)) + 
  stat_ecdf() +
  stat_function(fun = pnorm, 
                args = list(ET_xbar, ET_s), 
                color = "orange")


# qqplot
ggplot(geyser, aes(sample = erupt_time)) + 
  stat_qq(alpha = .5) +
  stat_qq_line(color = "orange")


# skewness and kutosis
skewK_ET <- tibble(statistic = c("Skewness", "Kurtosis", "Excess K"),
                theory = c(0, 3, 0),
                data   = round(c(skewness(geyser$erupt_time), 
                                 kurtosis(geyser$erupt_time), 
                                 kurtosis(geyser$erupt_time) - 3), 
                               2))
print.data.frame(skewK_ET, row.names = FALSE)
```

* Clearly bimodal

* Much bigger tilde in the data

* All plots show substantial deviation from the normal model (describe a little bit how, if you can)

* Difference between facet_wrap --> don't have to specify dimensions; facet_grid --> specifies dimensions

### *Creating another variable*

```{r creating a new variable}
# split the bimodal plot in half
geyser <- geyser %>% 
  mutate(erupt_category = ifelse(erupt_time < 3, "short", "long"))

# perfect normal distribution
ggplot(geyser, aes(x = erupt_time)) + 
  geom_density() +
  facet_wrap(~ erupt_category)


# empirical cdf
ggplot(geyser, aes(x = erupt_time)) + 
  stat_ecdf() +
  facet_wrap(~ erupt_category)


# qqplot
ggplot(geyser, aes(sample = erupt_time)) + 
  stat_qq(alpha = .5) +
  stat_qq_line(color = "orange") +
  facet_wrap(~ erupt_category)

```




***

#### **Bangladesh Wells**

```{r}
wells <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Bangladesh.csv")

```

```{r}
# Normality Assessment Here
ars_xbar <- mean(wells$Arsenic)
ars_s <- sd(wells$Arsenic)


# perfect normal distribution
ggplot(wells, aes(x = Arsenic)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(ars_xbar, ars_s),   
                color = "dark green") +
  scale_x_continuous(breaks = seq(ars_xbar - (4 * ars_s),  
                                  ars_xbar + (4 * ars_s), ars_s))


# empirical cdf
ggplot(wells, aes(x = Arsenic)) + 
  stat_ecdf() +
  stat_function(fun = pnorm, 
                args = list(ars_xbar, ars_s), 
                color = "dark green")


# qqplot
ggplot(wells, aes(sample = Arsenic)) + 
  stat_qq(alpha = .5) +
  stat_qq_line(color = "dark green")


# skewness and kutosis
skewK_ars <- tibble(statistic = c("Skewness", "Kurtosis", "Excess K"),
                theory = c(0, 3, 0),
                data   = round(c(skewness(wells$Arsenic), 
                                 kurtosis(wells$Arsenic), 
                                 kurtosis(wells$Arsenic) - 3), 
                               2))
print.data.frame(skewK_ars, row.names = FALSE)

```
* Data is right skewed and leptokurtic

* Only seeing the top half of the ecdf --> getting cut off right at inflection point

* QQ Plot shows that a right skewed distribution produces a concave up shape

***

#### **Semmelweiss**

```{r}
handwashing_start <- as.Date("1847-06-01")
data <- "https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/semmelweismonthly.csv"
by_month <- read_csv(file = url(data)) %>% 
  mutate(mortality = 100 * deaths / births,
         handwashing = ifelse(date < handwashing_start, "dirty", "washed"))
```

```{r}
# Normality Assessment Here
```

```{r}
# Normality Assessment by Groups Here, 
#before and after hand washing
```

***

#### **Oscar Ages**

```{r}
oscar <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oscarages.csv")
```

```{r}
# Normality Assessment Here
```

```{r}
# Normality Assessment by Groups Here
# win

```


***
```{r}
sessionInfo()
```
