---
title: "classwork"
author: "Elise Haylett"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, loading packages}
library(tidyverse)
```

```{r, mpg}
mileage <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv"))
glimpse(mileage)

```

```{r, old faithful}
oldfaithful <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv"))
glimpse(oldfaithful)

```

```{r, wells}
wells <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Bangladesh.csv"))
glimpse(wells)

```


Histogram
* modified bar plot
* cut the x-axis into equally sized chunks/intervals/"bins"
** there can be no gaps
** must cover the whole range of the data
* count how many data values are in each "bin"
```{r, histograms}
#uni modal data, follows a normal distribution
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram()

#bimodal
ggplot(oldfaithful, aes(x = erupt_time)) + 
  geom_histogram()

#extremely right skewed data
ggplot(wells, aes(x = Arsenic)) +
  geom_histogram()
```



```{r, density plots}
ggplot(mileage, aes(x = mpg)) + 
  geom_density()

ggplot(oldfaithful, aes(x = erupt_time)) + 
  geom_density()

ggplot(wells, aes(x = Arsenic)) + 
  geom_density()

```

```{r, density over a histogram}
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

ggplot(oldfaithful, aes(x = erupt_time)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

ggplot(wells, aes(x = Arsenic)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

```

```{r, density with color}
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram(aes(y = ..density..), bins = 10, color = "black") + 
  geom_density(fill = "salmon", color = "salmon", alpha = 0.50)

```

```{r, stat_ecdf}
ggplot(mileage, aes(x = mpg)) + 
  stat_ecdf()

ggplot(oldfaithful, aes(x = erupt_time)) + 
  stat_ecdf()

ggplot(wells, aes(x = Arsenic)) + 
  stat_ecdf()


```

#z-score
z-score is a unitless magnitude that will tell us distance

x-mu/sigma = zscore ---------------- mu = mean, sigma = standard deviation

x-mu -> distance of point from mean
sigma -> roughly the average distance of all points
*units will cancel each other out

# skewness and kurtosis
skewness - third moment
how you quantify skewness 
in perfectly symmetric data, the data will cancel each other out and the third moment/skewness = 0
kurotsis can only be positive
kurtosis of a bell curve is 3
kurtosis - fourth moment

anything that affects the mean and the standard deviation will affect the four moments

moment-based summaries in R
- mean()
- var()
- sd()
- skewness()
- kurtosis(), in the `moments` package in r

```{r, analysis on datasets}
(mean(mileage$mpg))
(sd(mileage$mpg))
# on average, points are spread out about 2.42 mpg above and below the mean

library(moments)
(skewness(mileage$mpg))
(kurtosis(mileage$mpg))
```

IQR = Q3 - Q1 --- range of the middle 50% of the data
Range = max - min
percentiles
- Q1 = 25th percentile/quantile
- median = 50th percentile/quantile
- Q3 = 75th percentile/quantile










