---
title: "classwork"
author: "Elise Haylett"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Classwork {.tabset .tabset-pills}

###Loading in the packages

```{r, loading packages}
library(tidyverse)
```

### Mileage Dataset
```{r, mpg}
mileage <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv"))
glimpse(mileage)

```

### Geyser Dataset
```{r, old faithful}
oldfaithful <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv"))
glimpse(oldfaithful)

```

### Wells Dataset
```{r, wells}
wells <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Bangladesh.csv"))
glimpse(wells)

```

### Histograms
Histogram
* modified bar plot
* cut the x-axis into equally sized chunks/intervals/"bins"
** there can be no gaps
** must cover the whole range of the data
* count how many data values are in each "bin"
```{r, histograms}
#uni modal data, follows a normal distribution
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram()

#bimodal
ggplot(oldfaithful, aes(x = erupt_time)) + 
  geom_histogram()

#extremely right skewed data
ggplot(wells, aes(x = Arsenic)) +
  geom_histogram()
```

### Density Plots

```{r, density plots}
ggplot(mileage, aes(x = mpg)) + 
  geom_density()

ggplot(oldfaithful, aes(x = erupt_time)) + 
  geom_density()

ggplot(wells, aes(x = Arsenic)) + 
  geom_density()

```

```{r, density over a histogram}
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

ggplot(oldfaithful, aes(x = erupt_time)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

ggplot(wells, aes(x = Arsenic)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density()

```

```{r, density with color}
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram(aes(y = ..density..), bins = 10, color = "black") + 
  geom_density(fill = "salmon", color = "salmon", alpha = 0.50)

```

### Stat_ecdf

```{r, stat_ecdf}
ggplot(mileage, aes(x = mpg)) + 
  stat_ecdf()

ggplot(oldfaithful, aes(x = erupt_time)) + 
  stat_ecdf()

ggplot(wells, aes(x = Arsenic)) + 
  stat_ecdf()


```


### Z-score

> Z-score
* Z-score is a unitless magnitude that will tell us distance

$$\frac{x - \mu}{\sigma} = z-score,$$
* $\mu$ = mean
* $\sigma$ = sd
* $x-\mu$ -> distance of point from mean
* $\sigma$ -> roughly the average distance of all points
  + note: units will cancel each other out



### Skewness and kurtosis

> Skewness and kurtosis

* skewness - third moment
* how you quantify skewness: 
** in perfectly symmetric data, the data will cancel each other out and the third moment/skewness = 0
* kurotsis can only be positive
* kurtosis of a bell curve is 3
* kurtosis - fourth moment

anything that affects the mean and the standard deviation will affect the four moments

* Moment-based summaries in R
    + mean()
    + var()
    + sd()
    + skewness()
    + kurtosis(), in the `moments` package in r

```{r, analysis on datasets}
(mean(mileage$mpg))
(sd(mileage$mpg))
# on average, points are spread out about 2.42 mpg above and below the mean

library(moments)
(skewness(mileage$mpg))
(kurtosis(mileage$mpg))
```




### 5 number summary

> 5-Number Summary

* minimum, Q1, median, Q3, maxium
* IQR = Q3 - Q1 --- range of the middle 50% of the data
* Range = max - min
* Percentiles
** Q1 = 25th percentile/quantile
** Median = 50th percentile/quantile
** Q3 = 75th percentile/quantile

```{r}
fivenum(mileage$mpg)
fivenum(mileage$mpg)[1]
range(mileage$mpg)
diff(range(mileage$mpg))

quantile(mileage$mpg, c(.25, .5, .75))
quantile(mileage$mpg, seq(.1, .9, .1)) 

#quartiles by 4 (.25, .5, .75)
#quintiles by 5 (.20, .40, .60, .80)
#deciles by 4 (.10, .20, .30, .40,..., .90)
```

```{r}
xbar <- mean(mileage$mpg)
s <- sd(mileage$mpg)

mileage <- mileage %>% 
  mutate(zscores = (mpg - xbar)/s,
         kpl = 0.43 * mpg,
         zscoresm = (kpl - mean(mileage$kpl))/sd(mileage$kpl)
  )
head(mileage)
```

```{r}
sort(mileage$mpg) #shows we only have two outliers: the min and the max
boxplot.stats(mileage$mpg) #output is a list since it has $ in front of names
boxplot.stats(sort(mileage$mpg)) #sorts in ascending order, only shows outliers

#plotting the graphs
boxplot(mileage$mpg)

ggplot(mileage, aes(y = mpg)) + 
  geom_boxplot() +
  coord_flip()
```

```{r}
wells %>% 
  summarize(lower_fence = fivenum(Arsenic)[2] - 1.5 * IQR(Arsenic),
            upper_fence = fivenum(Arsenic)[4] + 1.5 * IQR(Arsenic))

boxplot.stats(sort(wells$Arsenic))
            

ggplot(wells, aes(y = Arsenic)) +
  geom_boxplot() +
  labs(title = "arsenic levels from a sample of wells in bangledesh",
       y = "arsenic level in ppm") + 
  coord_flip()
```

```{r}
oldfaithful %>% 
  summarize(lower_fence = fivenum(erupt_time)[2] - 1.5 * IQR(erupt_time),
            upper_fence = fivenum(erupt_time)[4] + 1.5 * IQR(erupt_time))

boxplot.stats(sort(oldfaithful$erupt_time))
            

ggplot(oldfaithful, aes(y = erupt_time)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "old faithful eruption time of geyser",
       y = "eruption time in seconds") + 
  coord_flip()
```

```{r}
library(nycflights13)
library(ggplot2)
library(dplyr)

gapminder <- gapminder

ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, col = continent)) +
  geom_point(aes(size = pop)) + 
  scale_x_continuous("GDP per capita", limits = c(0, 50000)) +
  scale_y_continuous("Life expectancy", limits = c(40, 80))
  
```







