---
title: "R Notebook"
output: html_notebook
---

1) Exercise 3.1. You are welcome to do this by hand, but can you also figure out how to do it with R code? You do not have to show all the permutations in (b) but do report the differences in means.
```{r}
library(tidyverse)
library(combinat)
library(prob)

ctrl_mice <- c(5, 9)
exp_mice <- c(8, 10, 15)

# a
Diff <- exp_mice - ctrl_mice
observed <- mean(Diff)

# b
factorial(5)

# Replicate the entire data frame, permuting the promote variable
disc_perm <- disc %>%
  specify(promote ~ sex, success = "promoted") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5, type = "permute")

disc_perm %>%
  # Group by replicate
  group_by(replicate) %>%
  # Count per group
  count(promote)

disc_perm %>%
  # Calculate difference in proportion, male then female
  calculate(stat = "diff in props", order = c("male", "female"))
```


2) Exercise 3.2. Be sure to show us your thought processes and support your answers in this question, do not just answer "yes" and "no"?
```{r}
# I would be willing to tell my professor that she is wrong about the distribution of marbles (9,999 blue, 1 red) because it not is likely. The probability of pulling out 1 red marble out of 10,000 is .0001. It is more likely that 1 red marble can pulled out of a jar with a distribution of 1 red marble and 9 blue marbles. The probability of pulling out 1 red marble out of 10 is .1
```


3) Exercise 3.4. We subtract 1 from the number of simulation replications N (in the book's examples) and then add it back at the end. Why?
```{r}
#1 is added back because it includes the orinigal data as an extra resample. It avoids reporting an impossible p-value.
```


