---
title: "Think-Pair-Share: Traditional Two-Sample t-Test"
author: "Elise Haylett"
date: "4/19/2020"
output: html_document
---

```{r globaloptions, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

```{r loading packages and data files}
library(infer)
library(tidyverse)
```

```{r loading Semmelweiss data}
data <- "https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/semmelweisannualbyclinic.csv"
by_clinic <- read_csv(file = url(data)) %>% 
  mutate(mortality = 100 * deaths / births)

glimpse(by_clinic)
head(by_clinic)

handwashing_start <- as.Date("1847-06-01")
data <- "https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/semmelweismonthly.csv"
by_month <- read_csv(file = url(data)) %>% 
  mutate(mortality = 100 * deaths / births,
         handwashing = ifelse(date < handwashing_start, "dirty", "washed"))

glimpse(by_month)
head(by_month)
```

Consider the Semmelweiss activity from Week 1.  Conduct a traditional t-test to determine whether the average mortality for Clinic 1 was statistically significantly greater than the average mortality for Clinic 2 between 1841 and 1846 (one-tailed test). Report the test statistic (t) and the p-value. Were all of the necessary assumptions met here?

$$H_0: \mu_1 = \mu_2$$
$$H_A: \mu_1 > \mu_2$$
Using t-test function
```{r}
ggplot(by_clinic, aes(x = clinic, y = mortality)) +
 geom_boxplot()

ggplot(by_clinic, aes(sample = mortality)) +
  geom_qq() +
  geom_qq_line() +
  facet_grid(. ~ clinic)

t.test(mortality ~ clinic, data = by_clinic, alternative = "greater")
```

Using infer
```{r}
mortality <- by_clinic %>% 
  group_by(clinic) %>% 
  summarize(n = n(),
            average = mean(mortality))

diff_mortality <- diff(mortality$average)


n_replicates <- 1000
diff_mean_mortality <- by_clinic %>%
  specify(mortality ~ clinic) %>% 
  hypothesize(null = "independence") %>%  
  generate(reps = n_replicates, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("clinic 1", "clinic 2"))
  
diff_mean_mortality %>%
  filter(stat <= diff_mortality ) %>%
  summarize(p_val = n() / n_replicates)

```






Consider the Semmelweiss activity from Week 1. Conduct a traditional t-test to determine whether average mortality statistically significantly decreased after the introduction of handwashing in May 1847 (one-tailed test). Report the test statistic (t) and the p-value. Were all of the necessary assumptions met here?

$$H_0: \mu_{dirty} = \mu_{washed}$$
$$H_A: \mu_{dirty} > \mu_{washed}$$
```{r}
ggplot(by_month, aes(x = handwashing, y = mortality)) +
 geom_boxplot()

ggplot(by_month, aes(sample = mortality)) +
  geom_qq() +
  geom_qq_line() +
  facet_grid(. ~ handwashing)

handwashing <- by_month %>% 
  group_by(handwashing) %>% 
  summarize(n = n(),
            average = mean(mortality))

diff_handwashing <- diff(handwashing$average) # washed - dirty
```

Using t-test function
```{r}
t.test(mortality ~ handwashing, data = by_month, alternative = "greater")
```

Using infer
```{r}
n_replicates <- 1000
diff_mean_mortality <- by_month %>%
  specify(mortality ~ handwashing) %>% 
  hypothesize(null = "independence") %>%  
  generate(reps = n_replicates, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("dirty", "washed"))
  
diff_mean_mortality %>%
  filter(stat <= diff_handwashing ) %>%
  summarize(p_val = n() / n_replicates)

```
